---
title: "Visualization and EDA 2"
author: "Lina Marcinczyk"
date: "2026-10-06"
output: html_document
---

```{r}
library(tidyverse)
library(viridis)
library(patchwork)
library(p8105.datasets)

data("weather_df")
```

Scatterplot
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature(C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  )
```

Some other scales

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature(C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10 C", "0 C", "15 C")
  )+
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

Looking at color
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

themes

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

theme_bw() instead of greyscale
theme_minimal() takes the box off
theme_classic() takes the grid away
wes anderson theme package???

# A good reference plot!!

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  labs(
    title = "Seasonal Trends in Max Temp",
    x = "Date",
    y = "Max Temp",
    caption = "Max daily temp in three weather stations over the years",
    color = "Location"
  ) 
```

Two more weird useful plot things
Overlaying two dfs in different ways 

```{r}
central_park_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

Multiple panels with different plot types

```{r}
ggp_tmax_tmin =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

ggp_prcp_density = 
  weather_df |> 
  filter(prcp >0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5)

ggp_seasonal = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
  
(ggp_tmax_tmin + ggp_prcp_density)

(ggp_tmax_tmin / ggp_prcp_density)

(ggp_tmax_tmin + ggp_prcp_density) / ggp_seasonal
```
remove the legend from two of them for less bulk

### data manipulation
Start with factors

boxplots
```{r}
weather_df |> 
  mutate(name = fct_relevel(name, c("Molokai_HI", "CentralPark_NY", "Waterhole_WA"))) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

relevel you choose your order, reorder you choose by a factor

```{r}
weather_df |> 
  mutate(name = fct_reorder(name, tmax)) |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  labs(
    title = "Seasonal Trends in Max Temp",
    x = "Date",
    y = "Max Temp",
    caption = "Max daily temp in three weather stations over the years",
    color = "Location"
  ) 
```

Make the plot Jeff described

```{r}
weather_df |> 
  select(name, tmax, tmin) |> 
  pivot_longer(
    tmax:tmin, 
    names_to = "observation",
    values_to = "temp"
  ) |> 
  ggplot(aes(x = temp, fill = observation)) +
  geom_density(alpha = .5) +
  facet_grid(. ~ name)
```

```{r}
pups_df = 
  read_csv(
    "data/FAS_pups.csv", skip = 3, na = c("",".","NA")) |> 
  janitor::clean_names()

litters_df =
  read_csv(
    "data/FAS_litters.csv", na = c("",".","NA")) |> 
  janitor::clean_names() |> 
  separate(group, into = c("dose", "day_of_tx"), 3)

fas_df= 
  left_join(pups_df, litters_df, by = "litter_number")

fas_df |> 
  select(dose, day_of_tx, starts_with("pd")) |> 
  pivot_longer(
    starts_with("pd"),
    names_to = "outcome",
    values_to = "pn_day"
  ) |> 
  ggplot(aes(x = dose, y = pn_day)) +
  geom_boxplot() +
  facet_grid(day_of_tx ~ outcome)
```

