---
title: "Visualization and EDA 2"
author: "Lina Marcinczyk"
date: "2026-10-06"
output: html_document
---

```{r}
library(tidyverse)
library(viridis)
library(patchwork)
library(p8105.datasets)

data("weather_df")
```

Scatterplot
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature(C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  )
```

Some other scales

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  labs(
    title = "Temperature (Max vs Min)",
    x = "Max Temperature(C)",
    y = "Min Temperature (C)",
    color = "Location",
    caption = "Data from NOAA for three weather stations."
  ) +
  scale_x_continuous(
    breaks = c(-10, 0, 15),
    labels = c("-10 C", "0 C", "15 C")
  )+
  scale_y_continuous(
    trans = "sqrt",
    position = "right"
  )
```

Looking at color
```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location",
    discrete = TRUE
  )
```

themes

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme(legend.position = "bottom")
```

```{r}
weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point() +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")
```

theme_bw() instead of greyscale
theme_minimal() takes the box off
theme_classic() takes the grid away
wes anderson theme package???

# A good reference plot!!

```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  viridis::scale_color_viridis(
    name = "Location", 
    discrete = TRUE
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name) +
  labs(
    title = "Seasonal Trends in Max Temp",
    x = "Date",
    y = "Max Temp",
    caption = "Max daily temp in three weather stations over the years",
    color = "Location"
  ) 
```

Two more weird useful plot things
Overlaying two dfs in different ways 

```{r}
central_park_df =
  weather_df |> 
  filter(name == "CentralPark_NY")

molokai_df =
  weather_df |> 
  filter(name == "Molokai_HI")

ggplot(molokai_df, aes(x = date, y = tmax, color = name)) +
  geom_point() +
  geom_line(data = central_park_df)
```

Multiple panels with different plot types

```{r}
ggp_tmax_tmin =
  weather_df |> 
  ggplot(aes(x = tmin, y = tmax, color = name)) +
  geom_point()

ggp_prcp_density = 
  weather_df |> 
  filter(prcp >0) |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_density(alpha = .5)

ggp_seasonal = 
  weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point()
  
(ggp_tmax_tmin + ggp_prcp_density)

(ggp_tmax_tmin / ggp_prcp_density)

(ggp_tmax_tmin + ggp_prcp_density) / ggp_seasonal
```
remove the legend from two of them for less bulk


