---
title: "viz_and_eda_1"
author: "Lina Marcinczyk"
date: "2026-10-06"
output: html_document
---

```{r}
library(tidyverse)
library(ggridges)
library(hexbin)
```

```{r}
library(p8105.datasets)
data("weather_df")
```

# Basic scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```

# Advanced Scatterplot
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

# Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).

```{r}
weather_df |> 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE) 
```

# Odds and Ends
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex()
```

# Learning Assessment: In the preceding, we set the alpha aesthetic “by hand” instead of mapping it to a variable. This is possible for other aesthetics too. To check your understanding of this point, try to explain why the two lines below don’t produce the same result:



