---
title: "viz_and_eda_1"
author: "Lina Marcinczyk"
date: "2026-10-06"
output: html_document
---

```{r}
library(tidyverse)
library(ggridges)
library(hexbin)
```

```{r}
library(p8105.datasets)
data("weather_df")
```

# Basic scatterplot

```{r}
ggplot(weather_df, aes(x = tmin, y = tmax)) +
  geom_point()
```

```{r}
ggp_weather = 
  weather_df |>
  ggplot(aes(x = tmin, y = tmax)) 

ggp_weather + geom_point()
```

# Advanced Scatterplot
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_point(aes(size = prcp), alpha = .5) +
  geom_smooth(se = FALSE) +
  facet_grid(. ~ name)
```

# Learning Assessment: Write a code chain that starts with weather_df; focuses only on Central Park, converts temperatures to Fahrenheit, makes a scatterplot of min vs. max temperature, and overlays a linear regression line (using options in geom_smooth()).

```{r}
weather_df |> 
  mutate(
    tmax_fahr = tmax * (9 / 5) + 32,
    tmin_fahr = tmin * (9 / 5) + 32) |> 
  ggplot(aes(x = tmin_fahr, y = tmax_fahr)) +
  geom_point(alpha = .5) +
  geom_smooth(method = "lm", se = FALSE) 
```

# Odds and Ends
```{r}
weather_df |> 
  ggplot(aes(x = date, y = tmax, color = name)) +
  geom_smooth(se = FALSE)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = tmin)) +
  geom_hex()
```

# Learning Assessment: In the preceding, we set the alpha aesthetic “by hand” instead of mapping it to a variable. This is possible for other aesthetics too. To check your understanding of this point, try to explain why the two lines below don’t produce the same result:

```{r}
weather_df |> ggplot() + geom_point(aes(x = tmax, y = tmin), color = "blue")
weather_df |> ggplot() + geom_point(aes(x = tmax, y = tmin, color = "blue"))
```

# color is either inside or outside of the aes parentheses - outside actually produces blue, while inside produces orange?

# Solution: In the first attempt, we’re defining the color of the points by hand; in the second attempt, we’re implicitly creating a color variable that has the value blue everywhere; ggplot is then assigning colors according to this variable using the default color scheme.

# Univariate plots

#Histogram
```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_histogram(position = "dodge", binwidth = 2)
```

```{r}
weather_df |> 
  ggplot(aes(x = tmax, fill = name)) +
  geom_density(alpha = .4, adjust = .5, color = "blue") +
  geom_rug()
```

# boxplots
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_boxplot()
```


# Violin plots
```{r}
weather_df |> 
  ggplot(aes(x = name, y = tmax)) +
  geom_violin(aes(fill = name), alpha = .5) +
  stat_summary(fun = "median", color = "blue")
```

# Ridge plot

```{r}
weather_df |> 
  ggplot(aes(x = tmax, y = name)) + 
  geom_density_ridges(scale = .85)
```

# Learning Assessment: Make plots that compare precipitation across locations. Try a histogram, a density plot, a boxplot, a violin plot, and a ridgeplot; use aesthetic mappings to make your figure readable.

```{r}
weather_df |> 
  ggplot(aes(x = prcp, fill = name)) +
  geom_histogram()

weather_df |> 
  ggplot(aes(x = prcp)) +
  geom_density(aes(fill = name), alpha = .5)

weather_df |> 
  ggplot(aes(x = prcp, y = name)) +
  geom_density_ridges(scale = .85)

weather_df |> 
  ggplot(aes(x = prcp, y = name)) +
  geom_boxplot()

weather_df |> 
  filter(prcp > 0) |> 
  ggplot(aes(x = prcp, y = name)) + 
  geom_density_ridges(scale = .85)
```

# saving and embedding plots

```{r}

```






